---
title: "HW1"
author: ""
date: "2026-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

This is your first homework assignment. It focuses on plotting in the ways we already did in class. For the North Dakota data I would like you to:

1. Create a time series plot the *not seasonally adjusted* employment data.
2. Create a time series plot of both the *not seasonally adjusted* employment data and the seasonally adjusted employment data on the same chart.
3. Create a season plot of the *not seasonally adjusted* unemployment data.
4. Create a subseason plot, sometimes called a cycle plot for the *not seasonally adjusted* unemployment data. 

To aid you in these efforts I provide the following download and data preparation commands to allow you to worry about creating the necessary plots. 

```{r loading data}
lfdata <- read.csv("https://raw.githubusercontent.com/flynnecon/forecasting-2026/refs/heads/main/data/ndlf.csv")
lfts <- lfdata |> mutate(month=yearmonth(date)) |> select(-date) |> rename(date=month) |> as_tsibble(index=date) 
```

